---
title: post-processing
---

* need to have run "preprocessing.Rmd" first to generate files "sampPropChangeTimes.txt" and "ReChangeTimes.txt"

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/nadeaus/Repos/cov-swiss-phylogenetics/results_all/bdsky/log_files/NZTransmissionChains/")
knitr::opts_chunk$set(root.dir = "/Users/nadeaus/Repos/cov-swiss-phylogenetics/results_all/bdsky/log_files/NZTransmissionChains/")
```


```{r }
library(tidyverse)
library(lubridate)
library(ggdist)

source("../../../../scripts/figures_shared_vars.R")
system(command = "mkdir -p figures")
```

## Load data

```{r }
skyline_data <- NULL
for (clusters in c("min", "max")) {
    for (contact_tracing in c(TRUE, FALSE)) {
        for (sampUB in c("With sampling bound", "Without sampling bound" )) {
            pattern <- paste0("Re_skyline.", clusters, "_chains",
                              ".sampUB", ifelse(sampUB=="With sampling bound", "0.4", "1.0"),
                              ".", ifelse(contact_tracing, "1", "0"),
                              ".*.log")

            data <- NULL
            for (f in dir("results", pattern, full.names=TRUE)) {
                data <- bind_rows(data, read_tsv(f) %>% slice_tail(prop=0.9))
            }

            Re_temp_data <- data %>%
                pivot_longer(cols=starts_with("ReValues."),
                             names_to="interval", values_to="value",
                             names_prefix="ReValues.",
                             names_transform=list(interval=as.integer)) %>%
                select(Sample, interval, value) %>%
                mutate(variable="Re")

            skyline_temp_data <- bind_rows(Re_temp_data) %>%
                group_by(interval, variable) %>%
                summarize(median=median(value),
                          low=quantile(value, 0.025),
                          high=quantile(value, 0.975)) %>%
                mutate(clusters=clusters,
                       contact_tracing=contact_tracing,
                       sampUB=sampUB)

            skyline_data <- bind_rows(skyline_data, skyline_temp_data)
            
        }
    }
}
```

## Incoroporate interval dates

```{r }
finalSampleDate <- ymd("2020-11-30")
Re_dates <- read_csv("sequences/date_to_week.csv") %>%
    distinct(week) %>%
    filter(week<finalSampleDate) %>%
    arrange(desc(row_number())) %>%
    transmute(interval=row_number(), date=week+3.5)
```

## Load case count data

```{r}
LINE_LIST_RE_EST_LINK <- "https://raw.githubusercontent.com/covid-19-Re/dailyRe-Data/master/NZL-estimates.csv"
line_list_re_data <- read.csv(file = LINE_LIST_RE_EST_LINK) %>%
  filter(
    region == "NZL",
    data_type == "Confirmed cases", 
    estimate_type == "Cori_slidingWindow") %>%
  select(c(date, data_type, median_R_mean, median_R_highHPD, median_R_lowHPD)) %>%
  mutate(date = as.Date(date)) %>%
  rename(
    median = median_R_mean,
    yhigh = median_R_highHPD,
    ylow = median_R_lowHPD)
```

## Plot skylines

**In the final plots, make sure to add monthly/bi-weekly axis ticks.**

```{r }
ctChange <- ymd("2020-05-15")
color_values <- c("conf_case_counts" = "grey", "TRUE" = "#56B4E9", "FALSE" = "#F0E442")
color_labels <- c("conf_case_counts" = "Confirmed cases", 
                  "TRUE" = "Model with damping factor", 
                  "FALSE" = "Model without damping factor")

ggplot(skyline_data %>% filter(variable=="Re") %>% left_join(Re_dates),
       aes(date, median, fill = contact_tracing, color = contact_tracing)) +
    geom_ribbon(
        aes(ymin=low, ymax=high), alpha=0.5) +
    geom_ribbon(
        data = line_list_re_data, 
        aes(date, ymin=ylow, ymax=yhigh, fill = 'conf_case_counts', color = 'conf_case_counts'), alpha = 0.5) + 
    geom_line(alpha = 0.5) +
    geom_line(data = line_list_re_data, aes(color = 'conf_case_counts'), alpha = 0.5) +
    scale_color_manual(name = NULL, values = color_values, labels = color_labels) + 
    scale_fill_manual(name = NULL, values = color_values, labels = color_labels) + 
    geom_hline(yintercept=1, linetype="dashed") +
    facet_grid(rows=vars(clusters), cols=vars(sampUB), scales="free", labeller = labeller(clusters = chain_assumption_labs)) +
    scale_x_date(date_breaks="1 month", date_labels="%b", limits = c(as.Date("2020-01-01"), as.Date("2020-12-01"))) +
    labs(y = "Effective reproductive number", x = element_blank()) + 
    shared_theme
ggsave("figures/Re.pdf", width=15, height=5, units="in")
```

## Contact tracing effect size

```{r }
CT_data <- NULL
for (clusters in c("min", "max")) {
    for (contact_tracing in c(TRUE, FALSE)) {
        for (sampUB in c("With sampling bound", "Without sampling bound" )) {
            pattern <- paste0("Re_skyline.", clusters, "_chains",
                              ".sampUB", ifelse(sampUB=="With sampling bound", "0.4", "1.0"),
                              ".", ifelse(contact_tracing, "1", "0"),
                              ".*.log")

            data <- NULL
            for (f in dir("results", pattern, full.names=TRUE)) {
                data <- bind_rows(data, read_tsv(f) %>% slice_tail(prop=0.9))
            }

            CT_data <- bind_rows(CT_data,
                                 data %>%
                                 slice_tail(prop=0.9) %>%
                                 select(Sample, "CTFactor[0]", "CTFactor[1]") %>%
                                 pivot_longer(cols=2:3) %>%
                                 mutate(lessThanOne=(value<1.0)) %>%
                                 mutate(clusters=clusters,
                                        contact_tracing=contact_tracing,
                                        sampUB=sampUB))
        }
    }
}
```

## All-analyses damping factor figure for supplemental

```{r}
CT_data <- CT_data %>%
  mutate(season = case_when(
    name == "CTFactor[0]" ~ "After 15. May",
    name == "CTFactor[1]" ~ "Before 15. May"), 
  season = factor(season, levels = c("Before 15. May", "After 15. May")))

ggplot() +  
  geom_histogram(
    data = CT_data %>% filter(contact_tracing == "TRUE", clusters == "max"), 
    aes(x = value, fill = clusters), alpha = 0.5, bins = 30) + 
  geom_histogram(
    data = CT_data %>% filter(contact_tracing == "TRUE", clusters == "min"), 
    aes(x = value, fill = clusters), alpha = 0.5, bins = 30) + 
  facet_grid(season ~ sampUB) + 
  scale_fill_manual(values = c("#009E73", "#E69F00"), labels = chain_assumption_labs, name = "Data set conditioning") + 
  shared_theme + 
  labs(x = "Value of tranmission rate damping factor", y = "Posterior support")

ggsave("figures/contact_tracing_factor.pdf", width = double_col_width, height = single_col_width * 0.75, units = "cm")
```

## Write out variables for manuscript, copy figures into manuscript figure directory

```{r}
con <- file("../../../../manuscript/damping_factor_variables.tex", open = "a")

dampling_factor_medians <- CT_data %>%
  filter(contact_tracing == "TRUE") %>%
  group_by(sampUB, clusters, season) %>%
  summarize(median_dampling_factor = median(value), .groups = "drop")

before_15_may_max_dampling_percent_median_NZL_no_sampUB <- round(unlist(dampling_factor_medians %>%
  filter(sampUB == "Without sampling bound", clusters == "max", season == "Before 15. May") %>%
  select(median_dampling_factor)) * 100, digits = 0)

writeLines(text = paste0("\\newcommand\\before_15_may_max_dampling_percent_median_NZL_no_sampUB{", 
                         before_15_may_max_dampling_percent_median_NZL_no_sampUB, "}"), con = con)

before_15_may_min_dampling_percent_median_NZL_no_sampUB <- round(unlist(dampling_factor_medians %>%
  filter(sampUB == "Without sampling bound", clusters == "min", season == "Before 15. May") %>%
  select(median_dampling_factor)) * 100, digits = 0)

writeLines(text = paste0("\\newcommand\\before_15_may_min_dampling_percent_median_NZL_no_sampUB{", 
                         before_15_may_min_dampling_percent_median_NZL_no_sampUB, "}"), con = con)

close(con)

system("cp figures/contact_tracing_factor.pdf ../../../../figures/NZL_contact_tracing_factor.pdf")
system("cp figures/Re.pdf ../../../../figures/NZL_Re.pdf")
```

