---
title: "Plot Re estimates through time"
author: "Sarah Nadeau"
date: "30/04/2021"
output: html_document
---

Adapted from Tim's "postprocessing.Rmd" script.

```{r, setup, include=FALSE}
knitr::opts_knit$set(root.dir = "/Users/nadeaus/Repos/cov-swiss-phylogenetics")
knitr::opts_chunk$set(root.dir = "/Users/nadeaus/Repos/cov-swiss-phylogenetics")
```

```{r}
library(tidyverse)
library(lubridate)
source("scripts/figures_shared_vars.R")
source("../grapevine/utility_functions.R")
db_connection <- open_database_connection("server", config_file = "../grapevine/workdir/input/config.yml")
workdir <- "results_main"
```

## Load data

```{r, echo=F}
all_data <- NULL
for (clusters in c("min", "max")) {
    for (samples in c("all", "after May 1")) {
        for (contact_tracing in c(TRUE, FALSE)) {
            pattern <- paste0("Re_skyline.", clusters, "_chains",
                              ifelse(samples=="all", "", "_after_may_1"),
                              ".", ifelse(contact_tracing, "1", "0"),
                              ".*.log")

            data <- NULL
            for (f in dir(paste(workdir, "output/phylodynamics/results", sep = "/"), pattern, full.names=TRUE)) {
                data <- bind_rows(data, read_tsv(f) %>% slice_tail(prop=0.9))
            }

            re_data <- data %>%
                pivot_longer(cols=starts_with("ReValues."),
                             names_to="interval", values_to="value",
                             names_prefix="ReValues.",
                             names_transform=list(interval=as.integer)) %>%
                select(Sample, interval, value) %>%
                mutate(variable="Re")

            skyline_data <- re_data %>%
                group_by(interval, variable) %>%
                summarize(median=median(value),
                          low=quantile(value, 0.025),
                          high=quantile(value, 0.975)) %>%
                mutate(clusters=clusters,
                       samples=samples,
                       contact_tracing=contact_tracing)

            all_data <- bind_rows(all_data, skyline_data)
        }
    }
}
```

## Incoroporate Re interval dates

```{r }
dates <- tibble(interval=0:52) %>% mutate(date=ymd("2020-11-30")-3.5-interval*7)
```

## Load case count data

```{r}
LINE_LIST_RE_EST_LINK <- "https://raw.githubusercontent.com/covid-19-Re/dailyRe-Data/master/CHE-estimates.csv"
line_list_re_data <- read.csv(file = LINE_LIST_RE_EST_LINK) %>%
  filter(
    region == "CHE",
    data_type %in% c("Confirmed cases / tests", "Confirmed cases"), 
    estimate_type == "Cori_slidingWindow") %>%
  select(c(date, , data_type, median_R_mean, median_R_highHPD, median_R_lowHPD)) %>%
  mutate(date = as.Date(date)) %>%
  rename(
    median = median_R_mean,
    yhigh = median_R_highHPD,
    ylow = median_R_lowHPD)
```

## Plot all skylines

```{r }
re_all <- ggplot(all_data %>% left_join(dates) %>% filter(variable=="Re"),
       aes(date, median, col=contact_tracing, fill=contact_tracing)) +
  geom_ribbon(aes(x, ymin=ylow, ymax=yhigh),
              data=tibble(x=c(ymd("2020-06-15"), ymd("2020-09-30")),
                          ylow=c(0,0), yhigh=c(Inf,Inf),
                          median=0.5, contact_tracing=TRUE),
              fill="grey", col=NA, alpha=0.5) +
  geom_ribbon(aes(ymin=low, ymax=high), alpha=0.5) +
  geom_ribbon(
    data = line_list_re_data, 
    aes(date, ymin=ylow, ymax=yhigh, fill = data_type,
        col = data_type),
    alpha = 0.5) + 
  geom_line() +
  geom_hline(yintercept=1, linetype="dashed") +
  facet_grid(rows=vars(samples), cols=vars(clusters), scales="free") +
  scale_x_date(date_breaks="1 month", date_labels="%b %Y",
               limits = c(as.Date("2020-01-01"), as.Date("2020-12-01"))) +
  ylab("Re") + ggtitle("Re estimates") +
  theme(axis.text.x=element_text(angle=-45, hjust=0), legend.position = "bottom")

show(re_all)

ggsave(
  plot = re_all, 
  file = "figures/Re_with_conf_case_comparison.pdf", 
  width = double_col_width, 
  height = single_col_width, 
  units = "cm")
```

## Plot skyline w/ all data compared to case count Re

```{r }
re_methods_comparison <- ggplot(
  data = all_data %>% 
    left_join(dates) %>% 
    filter(
      variable=="Re", 
      samples == "all",
      !contact_tracing)) +
  geom_ribbon(
    aes(date, ymin=low, ymax=high, fill = clusters),
    alpha=0.5) +
  geom_ribbon(
    data = line_list_re_data, 
    aes(date, ymin=ylow, ymax=yhigh, fill = data_type),
    alpha = 0.5) + 
  geom_hline(yintercept=1, linetype="dashed") +
  scale_x_date(
    date_breaks="1 month", date_labels="%b %Y",
    limits = c(as.Date("2020-02-01"), as.Date("2020-12-01"))) +
  ylab("Re") +
  scale_fill_manual(
    name = element_blank(),
    labels = c(
      "max" = "Phylodynamics with\nsmallest plausible chains",
      "min" = "Phylodynamics with\nlargest plausible chains",
      "Confirmed cases" = "estimateR with\nconfirmed cases",
      "Confirmed cases / tests" = "estimateR with\nconfirmed cases / tests"),
    values = c("#E69F00", "#D55E00", "#56B4E9", "#0072B2")) +
  shared_theme + 
  theme(
    axis.text.x=element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.position = "bottom")

show(re_methods_comparison)

ggsave(
  plot = re_methods_comparison, 
  file = "figures/Re_all_data_with_conf_case_comparison.pdf", 
  width = double_col_width, 
  height = single_col_width * 0.75, 
  units = "cm")
```

## Plot case count Re only for intro slide

```{r }
re_case_counts <- ggplot() +
  geom_ribbon(
    data = line_list_re_data, 
    aes(date, ymin=ylow, ymax=yhigh, fill = data_type),
    alpha = 0.5) + 
  geom_hline(yintercept=1, linetype="dashed") +
  scale_x_date(
    date_breaks="1 month", date_labels="%b %Y",
    limits = c(as.Date("2020-02-01"), as.Date("2020-12-01"))) +
  ylab("Re") +
  scale_fill_manual(
    name = element_blank(),
    labels = c(
      "max" = "Phylodynamics with\nsmallest plausible chains",
      "min" = "Phylodynamics with\nlargest plausible chains",
      "Confirmed cases" = "estimateR with\nconfirmed cases",
      "Confirmed cases / tests" = "estimateR with\nconfirmed cases / tests"),
    values = c("#E69F00", "#D55E00", "#56B4E9", "#0072B2")) +
  shared_theme + 
  theme(
    axis.text.x=element_text(angle=45, hjust=1),
    axis.title.x = element_blank(),
    legend.position = "bottom")

show(re_case_counts)

ggsave(
  plot = re_case_counts, 
  file = "figures/Re_conf_cases.pdf", 
  width = double_col_width, 
  height = single_col_width * 0.75, 
  units = "cm")
```

## Plot datasets through time
The genome samples don't have an obvious delay compared to confirmed cases

```{r}
source("../grapevine/generate_figures/functions.R")
grapevine_results <- load_grapevine_results(
  workdir = "results_all/jan-dec_-01_max_sampling_1_travel_1_sim_context-sf_111_travel-wt")
samples_with_chain_sizes <- merge(
  x = grapevine_results$samples,
  y = grapevine_results$chains) %>%
  select(strain, date, chains_assumption, size)

sequence_dates <- samples_with_chain_sizes %>% 
  mutate(singleton = size == 1) %>%
  group_by(date, chains_assumption, singleton) %>%
  summarize(new_cases = n()) %>%
  mutate(type = "genome")
case_dates <- dplyr::tbl(db_connection, "ext_owid_global_cases") %>%
  filter(iso_country == "CHE") %>%
  select(date, new_cases) %>%
  collect() %>%
  mutate(type = "confirmed",
         chains_assumption = "Confirmed cases")

case_data <- merge(
  x = sequence_dates,
  y = case_dates,
  all = T) %>%
  mutate(period = case_when(
    date < as.Date("2020-04-01") ~ "Early period",
    T ~ "Later period"))

ggplot(
  data = case_data %>% filter(period == "Early period"),
  aes(x = date, y = new_cases)) + 
  facet_grid(chains_assumption ~ period, scales = "free", 
             labeller = labeller(chains_assumption = c(chain_assumption_labs, c("Confirmed cases" = "Confirmed cases")))) + 
  geom_col(aes(fill = singleton)) + 
  scale_x_date(
    date_breaks="1 week", date_labels="%b %d") +
  shared_theme + 
  theme(
    axis.text.x=element_text(angle=45, hjust=1),
    axis.title.x = element_blank()) + 
  labs(y = "Number of samples")

ggsave("figures/early_period_samples.pdf")
```

## Plot after-chain-sampled damping factor posteriors

```{r }
CT_data <- NULL
for (clusters in c("min", "max")) {
    for (samples in c("all", "after May 1")) {
        for (contact_tracing in c(TRUE, FALSE)) {
            pattern <- paste0("Re_skyline.", clusters, "_chains",
                              ifelse(samples=="all", "", "_after_may_1"),
                              ".", ifelse(contact_tracing, "1", "0"),
                              ".*.log")

            data <- NULL
            for (f in  dir(paste(workdir, "output/phylodynamics/results", sep = "/"), pattern, full.names=TRUE)) {
                data <- bind_rows(data, read_tsv(f) %>% slice_tail(prop=0.9))
            }

            CT_data <- bind_rows(CT_data,
                                 data %>%
                                  slice_tail(prop=0.9) %>%
                                  select(Sample, "CTFactor[0]", "CTFactor[1]", "CTFactor[2]") %>%
                                  pivot_longer(cols=2:4) %>%
                                  mutate(clusters=clusters,
                                         samples=samples,
                                         contact_tracing=contact_tracing))
        }
    }
}

ggplot(CT_data %>% filter(contact_tracing), aes(name, value, col=name, fill=name)) +
    geom_violin(scale="width") +
    facet_grid(rows=vars(samples), cols=vars(clusters), scales="free") +
    scale_x_discrete(labels=c("After Sep 30", "Jun 15 to Sep 30", "Before Jun 15")) +
    theme(axis.text.x=element_text(angle=-45, hjust=0)) +
    theme(legend.position = "none") +
    xlab("") + ylab("")
ggsave("figures/CT_effectSizes.pdf")
```

```{r}
data_to_plot <- CT_data %>% 
  mutate(
    period = factor(
      x = name,
      levels = c("CTFactor[2]", "CTFactor[1]", "CTFactor[0]"),
      labels = c("Before\nJun 15", "Jun 15\nto Sep 30", "After\nSep 30")),
    range = case_when(
      value == 1 ~ "No difference",
      value < 1 ~ "Re decrease")) %>%
  filter(contact_tracing, samples == "all")
data_labels <- data_to_plot %>%
  group_by(period, clusters) %>%
  mutate(n_samples = n()) %>%
  group_by(period, clusters, range) %>%
  summarize(posterior_prob = n() / n_samples[1],
            median_decrease = median(value))

ggplot() + 
  facet_grid(. ~ clusters, labeller = labeller(clusters = chain_assumption_labs)) + 
  geom_violin(
    data = data_to_plot %>% filter(range == "Re decrease"),
    aes(x = period, y = value),
    scale="width",
    fill = "grey",
    alpha = 0.5) + 
  geom_text(
    data = data_labels %>% filter(range == "Re decrease"),
    aes(
      x = period, y = 0, 
      label = paste("Inclusion\nprobability\n", signif(posterior_prob * 100, digits = 2), "%", sep = "")),
    size = 2,
    vjust = 0) + 
  labs(x = element_blank(), y = "f") + 
  shared_theme

ggsave(
  file = "figures/CT_all_data_effectSizes.pdf", 
  width = single_col_width, 
  height = single_col_width * 0.75, 
  units = "cm")
```
